---
title: "3rd homework assignment"
output: pdf_document
highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1 - Modern statistical methods (4 points)

First, load the data from the file `women.csv`. The data contains information about IQ and weight in $kg$ of 30
randomly selected women.


a) Compute the value of Pearson correlation coefficient for `IQ` and `weight`.
Compute the 95% confidence interval for the true correlation coefficient.
Round all the results to **three** decimal places.

| correlation coefficient estimate | lower bound for CI | upper bound for CI|
| ------------- |------------- |------------- |
| `insert` |`insert` |`insert` |

b) Use nonparametric bootstrap (perform 10 000 replications) to estimate 95% confidence interval (percentile) for the true correlation coefficient. Round the results to **three** decimal points.

| lower bound for bootstrap CI | upper bound for bootstrap CI|
|------------- |------------- |
|`insert` |`insert` |

c) Assume that the data come from a bivariate normal distribution.
Is there a connection between IQ and weight? Use the correlation coefficient to test it. Round corresponding p-value to **three** decimal points.

| p-value of the test | conclusion |
|-------------------| -----------------|
| `insert` | `insert` |


d) Use Monte Carlo simulations (perform 9 999 replications) to get a simulated p-value of the previous test. Round it to **three** decimal points.

| Simulated p-value |
|-------------------|
| `insert` |

\textcolor{blue}{\textbf{Hint:}} Assume normality of the data. Correlation coefficient does not depend on the value of means, nor variances of both variables. Hence they \textcolor{red}{\textbf{are not}} nuisance parameters and do not need to be taken into account.

\textcolor{red}{\textbf{Caution:}} Before every simulation run, do not forget to change `set.seed` of PRNG with your `UCO`:

```{r}
uco <- 235559  # insert your UCO
set.seed(uco)
```

\newpage
## Task 2 - Testing hypotheses (5 points)

Work with the data samples from `farm1.RData` and `farm2.RData` containing the weights of the total production (in `kg`) in different months at two farms. Your task is to compare their productions.

```{r}
load("farm1.RData")
load("farm2.RData")
```

Firstly answer the question, if the average weight of the production at the farm 1 equals 125 `kg` (as the owner of the farm proclaims), or less. Which statistical tool is appropriate (**explain** Your choice in details)? Why? What is the result (compute p-value and write the conclusion)?

| Name of the test | Explanation | p-value | conclusion |
| ----------------- | --------------------- | ------------ | ------------------ |
| `insert` | `insert` | `insert` | REJECTED / NOT REJECTED |

Now try to answer, if the average weights of production at the farms are the same. Which statistical test is appropriate (**explain** Your choice in details)? Why? What is the result (compute p-value and write the conclusion)? Support your conclusion by **one** suitable figure, which will visualize the results of your test (the averages of the weights, their difference etc.).

| Name of the test | Explanation | p-value | conclusion |
| ----------------- | ---------------------- | ------------- | ------------------ |
| `insert` | `insert` | `insert` | REJECTED / NOT REJECTED |

In each task \textcolor{red}{\textbf{check the assumptions}} of the tests you used and \textcolor{red}{\textbf{name them all}} in the `Explanation` sections.

## Task 3 - Linear model (6 points)

Work with the `cholesterol.RData` dataset. It contains information about the cholesterol levels, age, blood pressure, dietary preferences and smoking habits of 100 patients. Your task is to model the cholesterol levels. Try to create a model that best describes the cholesterol levels while being as simple as possible. 

| Your chosen model formula | adjusted R squared |
| ----------------- | ---------------------- |
| `insert e.g. cholesterol ~ age + smoker` | `insert` | 

\textcolor{blue}{\textbf{Hint:}} Consider different transformations of the explanatory variables at hand.

\textcolor{red}{\textbf{Warning:}} Keep in mind the limitations of linear models.

Interpret adjusted R squared. What does it mean?

| Value of adjusted R squared| Interpretation |
| ----------------- | ---------------------- |
| `insert value` | `insert text` | 

Interpret the coefficients of your chosen model. If your model uses a different number of coefficients as listed here, add or delete table rows.

| Variable | Value of coefficient | Interpretation |
| ----------------- | ---------------------- |---------------------- |
| `insert variable name` | `insert value` | `insert text` | 
| `insert variable name` | `insert value` | `insert text` | 
| `insert variable name` | `insert value` | `insert text` | 

Interpret the validity of your model (the F statistic in the models summary).

| Null hypothesis | p-value | Interpretation |
| ----------------- | ---------------------- | ---------------------- |
| `insert` | `insert` | `insert` | 
