---
title: "3rd homework assignment"
output: pdf_document
highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1 - Modern statistical methods (4 points)

First, load the data from the file `women.csv`. The data contains information about IQ and weight in $kg$ of 30 randomly selected women.

a)  Compute the value of Pearson correlation coefficient for `IQ` and `weight`. Compute the 95% confidence interval for the true correlation coefficient. Round all the results to **three** decimal places.

| correlation coefficient estimate | lower bound for CI | upper bound for CI |
|----------------------------------|--------------------|--------------------|
| -0.053                           | -0.406             | 0.313              |

b)  Use nonparametric bootstrap (perform 10 000 replications) to estimate 95% confidence interval (percentile) for the true correlation coefficient. Round the results to **three** decimal points.

| lower bound for bootstrap CI | upper bound for bootstrap CI |
|------------------------------|------------------------------|
| -0.443                       | 0.34                         |

c)  Assume that the data come from a bivariate normal distribution. Is there a connection between IQ and weight? Use the correlation coefficient to test it. Round corresponding p-value to **three** decimal points.

| p-value of the test | conclusion                                                                                                                   |
|-----------------|-------------------------------------------------------|
| 0.781               | I fail to reject the null hypothesis. There does not seem to be statistically significant correlation between IQ and weight. |

d)  Use Monte Carlo simulations (perform 9 999 replications) to get a simulated p-value of the previous test. Round it to **three** decimal points.

| Simulated p-value |
|-------------------|
| 0.779             |

\textcolor{blue}{\textbf{Hint:}} Assume normality of the data. Correlation coefficient does not depend on the value of means, nor variances of both variables. Hence they \textcolor{red}{\textbf{are not}} nuisance parameters and do not need to be taken into account.

\textcolor{red}{\textbf{Caution:}} Before every simulation run, do not forget to change `set.seed` of PRNG with your `UCO`:

```{r}
uco <- 492875  # insert your UCO
set.seed(uco)
```

\newpage

## Task 2 - Testing hypotheses (5 points)

Work with the data samples from `farm1.RData` and `farm2.RData` containing the weights of the total production (in `kg`) in different months at two farms. Your task is to compare their productions.

```{r}
load("farm1.RData")
load("farm2.RData")
```

Firstly answer the question, if the average weight of the production at the farm 1 equals 125 `kg` (as the owner of the farm proclaims), or less. Which statistical tool is appropriate (**explain** Your choice in details)? Why? What is the result (compute p-value and write the conclusion)?

| Name of the test  | Explanation                                                                                                                                                              | p-value | conclusion                                                                                                        |
|----------------|------------------|----------------|----------------------|
| one-sample t-test | We want to know if the samples come from a population with a specific mean (125kg). using Shappiro test, the data seem to be normally distributed so t-test can be used. | 0.18    | I fail to reject null hypothesis. There is not enough evidence to say that the average weight is less than 125kg. |

Now try to answer, if the average weights of production at the farms are the same. Which statistical test is appropriate (**explain** Your choice in details)? Why? What is the result (compute p-value and write the conclusion)? Support your conclusion by **one** suitable figure, which will visualize the results of your test (the averages of the weights, their difference etc.).

| Name of the test  | Explanation                                                                                                                                                                                | p-value | conclusion                                                                                                                                               |
|----------------|---------------------|----------------|--------------------|
| two-sample t-test | We have 2 independand datasets and want to test if their average product weight is the same. They seem to have normal distribution and similar variance, so two-sample t-test can be used. | 0.669   | I fail to reject null hypothesis. There is not enough evidence to suggest that the average weight of production between both farms differ significantly. |

In each task \textcolor{red}{\textbf{check the assumptions}} of the tests you used and \textcolor{red}{\textbf{name them all}} in the `Explanation` sections.

## Task 3 - Linear model (6 points)

Work with the `cholesterol.RData` dataset. It contains information about the cholesterol levels, age, blood pressure, dietary preferences and smoking habits of 100 patients. Your task is to model the cholesterol levels. Try to create a model that best describes the cholesterol levels while being as simple as possible.

| Your chosen model formula            | adjusted R squared |
|--------------------------------------|--------------------|
| log(cholesterol) \~ age + vegetarian | 0.967              |

\textcolor{blue}{\textbf{Hint:}} Consider different transformations of the explanatory variables at hand.

\textcolor{red}{\textbf{Warning:}} Keep in mind the limitations of linear models.

Interpret adjusted R squared. What does it mean?

| Value of adjusted R squared | Interpretation                                                                                                     |
|-------------------|-----------------------------------------------------|
| 0.967                       | It means that about 96.7% of the variation in log(cholesterol) (response variable) can be explained by this model. |

Interpret the coefficients of your chosen model. If your model uses a different number of coefficients as listed here, add or delete table rows.

| Variable             | Value of coefficient | Interpretation                                                                                                                                                                                |
|--------------------|---------------|--------------------------------------|
| age                  | 0.025                | Change in log(cholesterol) for each one-year increase in age. Can be interpreted also as 2.5% increase in log(cholesterol) for each aditional year of age.                                    |
| vegetarianmeat-eater | 0.121                | Difference between expected log(cholesterol) in vegetarians and meat-eaters. We would expect the log(cholesterol) to be on average about 0.121 lower for vegetarians compared to meat-eaters. |

Interpret the validity of your model (the F statistic in the models summary).

| Null hypothesis                                                                                                                                                               | p-value | Interpretation                                                                                                                                                                                     |
|------------------------|-----------------|--------------------------------|
| All of the regression coefficients are zero, which would imply that none of the predictors (age, being vegetarian) have any effect on the response variable log(cholesterol). | 2.2e-16 | I fail to reject the null hypothesis and can conclude that there is a strong evidence that at least one of the predictors is significantly associated with the response variable log(cholesterol). |
